Bootstrapping Multi-atlas Segmentation Using Multiple Automatically Generated Templates for the Segmentation of the Whole Hippocampus and Subfields

### 2014-01-27 paper.Rnw
Mallar's comments on Section 1 addressed

Also of interest to this manuscript are the methods that attempt to define
hippocampal subfields using standard T1- or T2-weighted data, of which there are
few.  \citet{VanLeemput2009} demonstrate that the applicability of hippocampal
subfield segmentation in T1-weighted images by Bayesian techniques using Markov
random field shape priors learned from 10 manual segmentations.  This work,
available as part of the \freesurfer{} package, is limited as the segmentation
omits the tail of the hippocampus and the protocol has yet to be fully
validated.  \citet{Yushkevich2009} manually segment hippocampal subfields on
high-resolution (either $0.2mm^3$ isotropic or $0.2mm \times 0.3mm \times 0.2mm$
resolution voxels) T2-weighted MR images acquired from five post-mortem medial
temporal lobe samples. Then, using nonlinear registration guided by shape-based
models of the subfield segmentations, and manually derived hippocampus masks of
the target images, the authors demonstrate accurate parcellation of hippocampal
subfields in clinical 3T T1-weighted MRI volumes.  Using \multiatlas{} with bias
correction techniques, \citet{Yushkevich2010} demonstrate a semi-automated
method of subfield segmentation on in vivo focal T2-weighted MR acquisitions of
the temporal lobe. Manual input is only needed to mark mark divisions between
the head, body and tail of the hippocampus on target images.

In this paper we describe a thorough validation of the \mb{} algorithm for the
fully automatic segmentation of the hippocampus and its subfields. First, we
address the feasibility of whole hippocampus segmentation with a limited number
of input atlases \citep{Chakravarty2013} by performing a multi-fold validation
experiment using a subset of the Alzheimer's Disease Neuroimaging Initiative
(ADNI) dataset manually segmented using the \prsnr{}-protocol, over a range of
atlas and template library sizes, registration and label fusion methods. This
experiment is then replicated with a first episode psychosis patient dataset
to determine how the behaviour of \mb{} changes when segmenting younger and
differently diseased subjects, and to ensure that we have not overfit our
parameters to the aging or neurodegenerative brain. A leave-one-out validation
of subfield segmentation is performed on the \citet{Winterburn2013} manually
segmented high-resolution MR atlases, using the optimal \mb{} parameters found
from the previous experiments. Finally, we validate our algorithm using
all of the data available in the \adnidataset{} sample and compare \mb{}
segmentations to other popular segmentation algorithms.

>>>>

Also of interest to this manuscript are the methods that attempt to define
hippocampal subfields using standard T1- or T2-weighted data, of which there are
few.  \citet{VanLeemput2009} demonstrate that the applicability of hippocampal
subfield segmentation in T1-weighted images by Bayesian techniques using Markov
random field shape priors learned from 10 manual segmentations.  This work,
available as part of the \freesurfer{} package, is limited as the segmentation
omits the tail of the hippocampus and the protocol has yet to be fully
validated.  \citet{Yushkevich2009} manually segment hippocampal subfields on
high-resolution (either $0.2mm^3$ isotropic or $0.2mm \times 0.3mm \times 0.2mm$
resolution voxels) T2-weighted MR images acquired from five post-mortem medial
temporal lobe samples. Then, using nonlinear registration guided by shape-based
models of the subfield segmentations, and manually derived hippocampus masks of
the target images, the authors demonstrate accurate parcellation of hippocampal
subfields in clinical 3T T1-weighted MRI volumes.  Using \multiatlas{} with bias
correction techniques, \citet{Yushkevich2010} demonstrate a semi-automated
method of subfield segmentation on in vivo focal T2-weighted MR acquisitions of
the temporal lobe. Manual input is only needed to mark divisions between
the head, body and tail of the hippocampus on target images.

In this paper we describe a thorough validation of the \mb{} algorithm for the
fully automatic segmentation of the hippocampus and its subfields. First, we
address the feasibility of whole hippocampus segmentation with a limited number
of input atlases \citep{Chakravarty2013} by performing a multi-fold validation
experiment using a subset of the Alzheimer's Disease Neuroimaging Initiative
(ADNI) dataset manually segmented using the \prsnr{}-protocol, over a range of
atlas and template library sizes, registration and label fusion methods. This
experiment is then replicated with a first episode psychosis patient dataset
to determine the behaviour of \mb{} when segmenting younger and differently
diseased subjects, and to ensure that we have not also overfit our parameters to
the aging or neurodegenerative brain. Next, we consider all of the images
available in the \adnidataset{} sample, and compare \mb{} segmentations to other
popular segmentation algorithms.  Finally, we conduct a leave-one-out validation
of subfield segmentation using the \citet{Winterburn2013} manually segmented
high-resolution MR atlases.
